Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-11 paper.tex
missing full stop

% working title, please change
\title{Python as a machine learning language:
       experiences from the \sklearn project}

@@

\begin{abstract}
The \sklearn project is an increasingly popular open source machine
learning library designed to be simple and efficient, accessible to non-experts,
and reusable in various contexts. In this paper, we present and discuss our
design choices for the application programming interface (API) of the project.
In particular, we describe the simple and limited interface shared by all
learning and processing units in the library and then discuss its advantages in
terms of composition and reusability. The paper also comments on implementation
details specific to the Python ecosystem and highlights obstacles faced by users
and developers of the library.
\end{abstract}

@@

\section{The \sklearn project}

@@

Started in 2009, \sklearn is developed by an international team
of several dozen volunteers, mostly researchers from various fields
(e.g., machine learning, natural language processing, or neuroscience).
The project also greatly benefits from many occasional
contributors proposing small code or documentation fixes or improvements.
Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
a platform which greatly facilitates this kind of collaboration.
Because of the large number of developers,
particular emphasis is put on keeping the project maintainable.
In particular, code must follow specific
quality guidelines, such as style consistency and unit-test coverage. In
addition, complete documentation and examples are required for all
features.
Finally, additions to the code base must pass code review
by at least two developers not involved in the implementation
of the proposed addition.

@@

In this paper, we describe and motivate the design choices used in
building the library. First, in section \ref{sec:api}, we give the
central application programming interface (API). Then, in section
\ref{sec:implementation}, we describe the implementation trade-offs.
In section \ref{sec:weaknesses}, we analyze the limitations of the
technologies we rely on; and finally, in section \ref{sec:comparison}, we
position \emph{scikit-learn} with respect to other major projects.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator. Through duck typing, composed
estimators can be used wherever usual estimators can be used. In particular,
they integrate within model selection routines, allowing one to optimize at once
every step of a complex workflow. Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects or in a parallel
fashion through \texttt{FeatureUnion} objects.

@@

As an example, the following snippet illustrates how to create a complex worflow
computing both linear PCA and kernel PCA components on \texttt{X\_train} (through a
\texttt{FeatureUnion}), selecting the best of those features and then feeding
them to a logistic regression model:
\begin{verbatim}
Pipeline([("pca", FeatureUnion([('linear_pca', PCA()),
                                ('kernel_pca', KernelPCA())])),
          ("feature_selection", SelectKBest()),
          ("logistic_model", LogisticRegression())
]).fit(X_train, y_train).predict(X_test)
\end{verbatim}

>>>>

% working title, please change
\title{Python as a machine learning language:
       experiences from the scikit-learn project}

@@

\begin{abstract}
The \sklearn project is an increasingly popular open source machine
learning library designed to be simple and efficient, accessible to non-experts,
and reusable in various contexts. In this paper, we present and discuss our
design choices for the application programming interface (API) of the project.
In particular, we describe the simple and elegant interface shared by all
learning and processing units in the library and then discuss its advantages in
terms of composition and reusability. The paper also comments on implementation
details specific to the Python ecosystem and analyzes obstacles faced by users
and developers of the library.
\end{abstract}

@@

\section{The scikit-learn project}

@@

Started in 2009, \sklearn is developed by an international team
of a dozen of core developers, mostly researchers from various fields
(e.g., machine learning, natural language processing, or neuroscience).
The project also greatly benefits from many occasional
contributors proposing small code or documentation fixes or improvements.
Development proceeds on GitHub (\url{https://github.com/scikit-learn}),
a platform which greatly facilitates this kind of collaboration.
Because of the large number of developers,
particular emphasis is put on keeping the project maintainable.
In particular, code must follow specific
quality guidelines, such as style consistency and unit-test coverage. In
addition, complete documentation and examples are required for all
features.
Finally, additions to the code base must pass code review
by at least two developers not involved in the implementation
of the proposed addition.

@@

In a previous paper \citep{pedregosa2011}, \sklearn was briefly presented and
benchmarked against several competitors. In this paper, we instead focus on an
in-depth discussion and analysis of the design choices used in
building the library. First, in section \ref{sec:api}, we present the
central application programming interface (API). Then, in section
\ref{sec:implementation}, we describe the implementation trade-offs.
In section \ref{sec:weaknesses}, we analyze the limitations of the
technologies we rely on; and finally, in section \ref{sec:comparison}, we
compare \emph{scikit-learn} to other major projects in terms of API.

@@

A distinguishing feature of the \sklearn API is its ability to
compose new estimators from several base estimators. Composition mechanisms can
be used to combine typical machine learning workflows into a single object which
is itself an estimator. Thanks to duck typing, composed
estimators can be used wherever usual estimators can be used. In particular,
they integrate within model selection routines, allowing one to optimize at once
every step of a complex workflow. Composition of estimators can be done in two
ways: either sequentially through \texttt{Pipeline} objects or in a parallel
fashion through \texttt{FeatureUnion} objects.

@@

As an example, the following snippet illustrates how to create a complex worflow
computing both linear PCA and kernel PCA components on \texttt{X\_train}
(through a \texttt{FeatureUnion}), selecting the 100 best of those features and
then feeding them to an $\ell_2$-regularized logistic regression model:
\begin{verbatim}
from sklearn.pipeline import Pipeline
from sklearn.decomposition import PCA, KernelPCA
from sklearn.feature_selection import SelectKBest

Pipeline([("pca", FeatureUnion([('pca', PCA()),
                                ('kpca', KernelPCA(kernel="rbf"))])),
          ("feature_selection", SelectKBest(k=100)),
          ("logistic_model", LogisticRegression(penalty="l2"))
]).fit(X_train, y_train).predict(X_test)
\end{verbatim}
